Capstone Movielens Report

### 2025-06-01 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
Model training with default parameters described.

### UMGYDE Model: Support Functions
\

@@

### UMGYDE Model Building
\

::: {.noteblock data-latex=""}
The complete source code of building and training the current model is available in the [UMGYDE Model (UMGYDEM) Building](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1745) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1745) script on _GitHub_.
:::

Below, we provide the most significant line of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
```
```{r}
```

::: {.noteblock data-latex=""}
We use the [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv) function described above to compute the _UMGYD Effect_.
:::

We can now construct predictors and calculate the _RMSE_ of the current model using the [calc_UMGYD_effect_RMSE.cv](#func.calc_UMGYD_effect_RMSE.cv) function described above:

```{r, eval=FALSE}
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UMGYDE)
```

@@

### UMGYDE Model Tuning
\

@@

>>>>

### UMGYDE Model: Helper Functions
\

@@

### UMGYDE Model Building With Default Parameters
\

::: {.noteblock data-latex=""}
The complete source code of building and training the current model is available in the [UMGYDEM Training with default parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1745) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1745) script on _GitHub_.
:::

We are going to use our helper function [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) described in the [UMGYDE Model: Helper Functions] section, which, as we explained, internally calls the R function [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess).

We will start by calling the `train_UMGY_SmoothedDay_effect.cv` with default parameters, which means that the `loess` function will be implicitly called with `degree = 1` and `span = 0.75` parameters (the default values for our custom [loess_de](#func.loess_de) function, which, in turn, calls the `loess` under the hood).

Below, we provide the most significant line of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
  cv.UMGYDE.default_params <- train_UMGY_SmoothedDay_effect.cv()
```
```{r}
str(cv.UMGYDE.default_params)
```

Below is the visual representation of the UMGYD Effect data we have just computed:

```{r}
cv.UMGYDE.default_params |>
  ggplot(aes(x = days)) +
  geom_point(aes(y = de), size = 3, alpha = .5, color = "grey") + 
  geom_line(aes(y = de_smoothed), color = "red")
```

Let's calculate the _RMSE_ of the current model using the [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) function described above:

```{r, eval=FALSE}
cv.UMGYDE.default_params.RMSE <- cv.UMGYDE.default_params |>
  calc_UMGY_SmoothedDay_effect.RMSE.cv()

RMSEs.ResultTibble.UMGYDE <- RMSEs.ResultTibble.rglr.UMGYE |> 
  RMSEs.AddRow("UMGYDE (Default) Model", 
               cv.UMGYDE.default_params.RMSE,
               comment = "User+Movie+Genre+Year+Day Effect (UMGYDE) Model 
computed using `stats::loess` function with `degree=1` & `span=0.75` parameter values.")
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UMGYDE)
```

Unfortunately, with the default parameters, we obtained an even worse result than before.
Let's tune the current model by adjusting the `degree` and `span` parameters and see what we get.

::: {.noteblock data-latex=""}
Since the `degree` parameter can only take three discrete values, 0, 1, and 2, we will perform the tuning in three steps - one per each possible value of the `degree` parameter.
:::

@@

### UMGYDE Model Tuning by `degree` and `span` Parameters
\

@@

#### UMGYDE Model Tuning: Step 1 (`degree = 0`)
\

::: {.noteblock data-latex=""}
The complete source code of the ***Step 1*** of the current model tuning is available in the [UMGYDE Model Tuning: Step 1 (`degree = `)](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1823) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1823) script on _GitHub_.
:::

#### UMGYDE Model Tuning: Step 2 (`degree = 1`)
\

::: {.noteblock data-latex=""}
The complete source code the ***Step 2*** of the current model tuning is available in the [UMGYDE Model Tuning: Step  (`degree = `)](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1) script on _GitHub_.
:::

#### UMGYDE Model Tuning: Step 3 (`degree = 2`)
\

::: {.noteblock data-latex=""}
The complete source code the ***Step 3*** of the current model tuning is available in the [UMGYDE Model Tuning: Step  (`degree = `)](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1) script on _GitHub_.
:::
